MSSM-Penn-Pitt - ACC and DLPFC - DREAM - FEATURECOUNTS - GRCh38 - CQN - Sparse Model

### 2020-03-02 cmc-rnaseq-de/MSSM-Penn-Pitt_ACC_DLPFC_FEATURECOUNTS_GRCh38_CQN_Sparse.Rmd
fix tissue correlation plot, fix contrasts

### Tissue Correlation
```{r tissue_correlation}
hasReps = COVARIATES %>%
  dplyr::select(Individual_ID) %>% 
  table

df_dlpfc = COVARIATES %>% 
  filter( Individual_ID %in% names(hasReps[hasReps==2])) %>%
  filter(grepl("DLPFC", Dx.Tissue)) %>% 
  dplyr::select(SampleID, Individual_ID, Dx.Tissue)
df_acc = COVARIATES %>%
  filter( Individual_ID %in% names(hasReps[hasReps==2])) %>%
  filter(grepl("ACC", Dx.Tissue)) %>% 
  dplyr::select(SampleID, Individual_ID, Dx.Tissue)

@@

@@

df_pca = merge(pca$x, as.data.frame(cov_dream), by.x="row.names", by.y="SampleID" )

@@

### Coexpression of residual expression 
```{r coexp2, cache=FALSE, fig.height=5, fig.width=5}
tmp = RESIDUAL.GENE_EXPRESSION
tmp[is.na(tmp)] = 0
cr = cor(t(tmp))
hist(cr, main = 'Distribution of correlation between genes', xlab = 'Correlation')
```

@@

```{r decompse.normalise.data2.1, fig.height=8, fig.width=12, results='asis'}
# Eucledian tree based analysis
COVARIATES.tmp = data.matrix(COVARIATES[,c('Dx.Tissue','Institution')])
COVARIATES.tmp[is.na(COVARIATES.tmp)] = 0

@@

### Differential expression analysis
Genes that are differentially expressed at an FDR <= 0.05 are
```{r diffExp, fig.height=10, fig.width=15}

formula <- glue::glue("~ 0 + ", glue::glue_collapse(adjust.covars, sep = " + ")) %>% 
  glue::glue(" + (1|{random_effect})")

L1 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
                                   formula = formula, 
                                   data = COVARIATES, 
                                   coefficient = c("Dx.TissueOther_ACC", "Dx.TissueControl_ACC"))
L2 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
                                   formula = formula, 
                                   data = COVARIATES, 
                                   coefficient = c("Dx.TissueSCZ_ACC", "Dx.TissueControl_ACC"))
L3 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
                                   formula = formula, 
                                   data = COVARIATES, 
                                   coefficient = c("Dx.TissueSCZ_ACC", "Dx.TissueOther_ACC"))
L4 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
                                   formula = formula, 
                                   data = COVARIATES, 
                                   coefficient = c("Dx.TissueOther_DLPFC", "Dx.TissueControl_DLPFC"))
L5 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
                                   formula = formula, 
                                   data = COVARIATES, 
                                   coefficient = c("Dx.TissueSCZ_DLPFC", "Dx.TissueControl_DLPFC"))

L6 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
                                   formula = formula, 
                                   data = COVARIATES, 
                                   coefficient = c("Dx.TissueSCZ_DLPFC", "Dx.TissueOther_DLPFC"))
L = cbind(L1, L2, L3, L4, L5, L6)

contrast <- c(glue::glue_collapse(names(L1[L1 != 0]), "-"), 
              glue::glue_collapse(names(L2[L2 != 0]), "-"),
              glue::glue_collapse(names(L3[L3 != 0]), "-"),
              glue::glue_collapse(names(L4[L4 != 0]), "-"),
              glue::glue_collapse(names(L5[L5 != 0]), "-"),
              glue::glue_collapse(names(L6[L6 != 0]), "-"))

@@

# Get differential expression
DE = lapply(glue::glue("L{1:6}"), function(i, FIT){
  topTable(FIT, coef=i, number = Inf) %>%
    rownameToFirstColumn('gene_id') %>% 
    left_join(backgroundGenes)
}, FIT.CONTR) 
names(DE) <- contrast

>>>>

### Tissue Correlation
```{r tissue_correlation}
hasReps = rownameToFirstColumn(COVARIATES, "SampleID") %>%
  dplyr::select(Individual_ID) %>% 
  table

df_dlpfc = rownameToFirstColumn(COVARIATES, "SampleID") %>% 
  filter( Individual_ID %in% names(hasReps[hasReps==2])) %>%
  filter(grepl("DLPFC", Dx.Tissue)) %>% 
  dplyr::select(SampleID, Individual_ID, Dx.Tissue)
df_acc = rownameToFirstColumn(COVARIATES, "SampleID") %>%
  filter( Individual_ID %in% names(hasReps[hasReps==2])) %>%
  filter(grepl("ACC", Dx.Tissue)) %>% 
  dplyr::select(SampleID, Individual_ID, Dx.Tissue)

@@

res_std <- rownameToFirstColumn(RESIDUAL.GENE_EXPRESSION, "ensembl_gene_id")

@@

df_pca = merge(pca$x, as.data.frame(rownameToFirstColumn(COVARIATES, "SampleID")), by.x="row.names", by.y="SampleID" ) %>% 
  tidyr::separate(Dx.Tissue, into = c("Dx", "Tissue"), sep = "_")

@@

### Coexpression of residual expression 
```{r coexp2, cache=FALSE, fig.height=5, fig.width=5}
cr = cor(t(RESIDUAL.GENE_EXPRESSION))
hist(cr, main = 'Distribution of correlation between genes', xlab = 'Correlation')
```

@@

```{r decompse.normalise.data2.1, fig.height=8, fig.width=12, results='asis'}
# Eucledian tree based analysis
COVARIATES.tmp = data.matrix(COVARIATES[,c('Dx.Tissue','Institution')])

@@

### Differential expression analysis
Genes that are differentially expressed at an FDR <= 0.05 are:
```{r set-formula}
formula <- glue::glue("~ 0 + ", glue::glue_collapse(adjust.covars, sep = " + ")) %>% 
  glue::glue(" + (1|{random_effect})")
```
`r formula`
```{r diffExp, fig.height=10, fig.width=15}
L1 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
                                   formula = formula, 
                                   data = COVARIATES, 
                                   coefficient = c("Dx.TissueOther_ACC", "Dx.TissueControl_ACC"))
L2 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
                                   formula = formula, 
                                   data = COVARIATES, 
                                   coefficient = c("Dx.TissueSCZ_ACC", "Dx.TissueControl_ACC"))
L3 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
                                   formula = formula, 
                                   data = COVARIATES, 
                                   coefficient = c("Dx.TissueSCZ_ACC", "Dx.TissueOther_ACC"))
L4 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
                                   formula = formula, 
                                   data = COVARIATES, 
                                   coefficient = c("Dx.TissueOther_DLPFC", "Dx.TissueControl_DLPFC"))
L5 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
                                   formula = formula, 
                                   data = COVARIATES, 
                                   coefficient = c("Dx.TissueSCZ_DLPFC", "Dx.TissueControl_DLPFC"))
L6 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
                                   formula = formula, 
                                   data = COVARIATES, 
                                   coefficient = c("Dx.TissueSCZ_DLPFC", "Dx.TissueOther_DLPFC"))
L7 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
                                   formula = formula, 
                                   data = COVARIATES, 
                                   coefficient = c("Dx.TissueSCZ_DLPFC", "Dx.TissueSCZ_ACC"))
L8 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
                                   formula = formula, 
                                   data = COVARIATES, 
                                   coefficient = c("Dx.TissueControl_DLPFC", "Dx.TissueControl_ACC"))
L9 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
                                   formula = formula, 
                                   data = COVARIATES, 
                                   coefficient = c("Dx.TissueOther_DLPFC", "Dx.TissueOther_ACC"))

L = cbind(L1, L2, L3, L4, L5, L6, L7, L8, L9)

contrast <- c(glue::glue_collapse(names(L1[L1 != 0]), "-"), 
              glue::glue_collapse(names(L2[L2 != 0]), "-"),
              glue::glue_collapse(names(L3[L3 != 0]), "-"),
              glue::glue_collapse(names(L4[L4 != 0]), "-"),
              glue::glue_collapse(names(L5[L5 != 0]), "-"),
              glue::glue_collapse(names(L6[L6 != 0]), "-"),
              glue::glue_collapse(names(L7[L7 != 0]), "-"),
              glue::glue_collapse(names(L8[L8 != 0]), "-"),
              glue::glue_collapse(names(L9[L9 != 0]), "-")
)

@@

# Get differential expression
DE = lapply(glue::glue("L{1:9}"), function(i, FIT){
  topTable(FIT, coef=i, number = Inf) %>%
    rownameToFirstColumn('gene_id') %>% 
    left_join(backgroundGenes)
}, FIT.CONTR) 
names(DE) <- contrast
